Optimal (Multiway) Spatial Joins

### 2024-05-23 msj.tex
yufei

\def\out{\mathrm{OUT}}

@@

\begin{figure}
    \includegraphics[height=25mm]{./artwork/four-segs}
    \figcapup
    \caption{Constructing $r_1'$ from $r_1$}
    \label{fig:four-segs}
    \figcapdown
\end{figure}

\noindent {\bf Step 1.} We will accomplish this step by solving an instance of $(k-2)$-SJ. To construct the instance, for each rectangle $r_1 \in R_1$, we compute four segments:
\myitems{
    \item $h_1$: the lowest segment (i.e., with the smallest y-coordinate) among all the segments in $H$ that cross $r$;
    \item $h_2$: the highest segment (i.e., with the largest y-coordinate) among all the segments in $H$ that cross $r$;
    \item $v_1$: the leftmost segment (i.e., with the smallest x-coordinate) among all the segments in $V$ that cross $r$;
    \item $v_2$: the rightmost segment (i.e., with the largest x-coordinate) among all the segments in $V$ that cross $r$.
}
Define $r_1'$ as the rectangle defined by the above four segments in the way illustrated in Figure~\ref{fig:four-segs}, where $r_1'$ is the shaded area. Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$. Define
\myeqn{
    R_1' &=& \set{r_1' \mid r_1 \in R_1 \text{ and its $r_1'$ defined as above exists}}. \nn
}
Computing the corresponding ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\EuScript{A}$. By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$. It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time and $O(n)$ space.

@@

\extraspacing {\bf Step 2.}

>>>>

\def\cross{\mathrm{cross}}
\def\out{\mathrm{OUT}}

@@

\noindent {\bf Step 1.} We will accomplish this step by solving an instance of $(k-2)$-SJ. To construct the instance, for each rectangle $r_1 \in R_1$, we compute four segments:
\myitems{
    \item $h_1$: the lowest segment (i.e., with the smallest y-coordinate) among all the segments in $H$ that cross $r$;
    \item $h_2$: the highest segment (i.e., with the largest y-coordinate) among all the segments in $H$ that cross $r$;
    \item $v_1$: the leftmost segment (i.e., with the smallest x-coordinate) among all the segments in $V$ that cross $r$;
    \item $v_2$: the rightmost segment (i.e., with the largest x-coordinate) among all the segments in $V$ that cross $r$.
}
Define $r_1'$ as the rectangle defined by the above four segments in the way illustrated in Figure~\ref{fig:type1}(a), where $r_1'$ is the shaded area. Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$. Define
\myeqn{
    R_1' &=& \set{r_1' \mid r_1 \in R_1 \text{ and its $r_1'$ defined as above exists}}. \nn
}
Computing the corresponding ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\EuScript{A}$. By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$. It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time and $O(n)$ space.

@@

\extraspacing {\bf Step 2.} We are ready to explain how to produce $\J_1(R_1, ..., R_{k-2}, H, V)$ from $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$. Our algorithm for this purpose will find a set $S$ satisfying
\myeqn{
    \J_1(R_1, ..., R_{k-2}, H, V) 
    \subseteq 
    S
    \subseteq 
    \J(R_1, ..., R_{k-2}, H, V). \nn
}
Recall that $\J(R_1, ..., R_{k-2}, H, V)$ is the result of the original H-V $k$-SJ problem, which has a size of $\out$. Only the $k$-tuples in $S$ that are of Type 1 will be reported. We will achieve the purpose in time $O(n \log n + |S|)$ and space $O(n + |S|)$.

Let us now fix an arbitrary $(k-2)$-tuple in $(r_1, ..., r_{k-2}) \in \Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$. Denote by 
\myeqn{
    B = \cap_{i=1}^{k-2} r_i. \nn
}
Note that $B$ is a rectangle itself. We say that $B$ is 
\myitems{
    \item {\em left-bounded} by $r_i$, if the left edge of $B$ is a part of the left boundary of $r_i$ for some $i \in [k-2]$;
    \item {\em bottom-bounded} by $r_j$, if the bottom edge of $B$ is a part of the bottom boundary of $r_j$ for some $j \in [k-2]$.
}
Figure~\ref{fig:type1}(b) shows an example with rectangles $r_1, r_2$, and $r_3$ (here $k = 5$). Their intersection is the gray rectangle $B$, which is left-bounded by $r_3$ and bottom-bounded by $r_2$.

We will compute:
\myitems{
    \item $\cross(B, H)$, the set of segments in $H$ that cross $B$; 
    \item $\cross(B, V)$, the set of segments in $V$ that cross $B$.
}

\begin{figure} 
    \begin{tabular}{cc}
            \includegraphics[height=25mm]{./artwork/four-segs} &
            \includegraphics[height=28mm]{./artwork/oracle} \\ 
            (a) Constructing $r_1'$ from $r_1$ & 
            (b) Crossing oracles
    \end{tabular}
    \figcapup
    \caption{Illustrations for Section~\ref{sec:hv:type1}}
    \label{fig:type1}
    \figcapdown
\end{figure}

We say that a $k$-tuple in $\J(R_1, ..., R_{k-2}, H, V)$ is {\em supported} by $(r_1, ..., r_{k-2})$ if it has the form $(r_1, ..., r_{k-2}, h, v)$, where $h \in H$ and $v \in V$. This means that the intersection of

Let us discuss how to generate all the $k$-tuples in
